Social Conditions of Fear

### 2023-11-22 scripts/manuscript.qmd
attempting to deal with missing values

library(tidyverse)
library(dplyr)
library(tidyr)
library(here)
library(rio)
library(janitor)
library(knitr)
library(kableExtra)
```

@@

gss <- import(here("data", "GSS.xlsx")) %>%
  as_tibble() %>% 
  characterize() %>% 
  clean_names()

@@

```{r}
#| label: select variables
data <- gss2022 %>% 
    select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racerank1", "uscitzn", "dwelown")

@@

```{r}
#| label: recode variables
tidydata <- data %>% 
  mutate(fear = 
           case_when( fear == ".i:  Inapplicable" | fear == ".n: No answer" | 
                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ "missing", 
                      fear == "YES" ~ "yes", 
                      fear == "NO" ~ "no")) %>% 
  filter(fear == "yes" | fear == "no") %>% 
  mutate(racecombo = 
           case_when(racerank1 == "White" ~ "White",
                     racerank1 == "Black or African American" ~ "Black",
                     racerank1 == "Hispanic" ~ "Hispanic",
                     racerank1 == "Asian Indian" | racerank1 ==  "Chinese" |
                       racerank1 ==  "Filipino" | racerank1 ==  "Japanese" | 
                       racerank1 ==  "Korean" |  racerank1 ==  "Vietnamese" |
                       racerank1 ==  "Other Asian" ~ "Asian",
                     racerank1 == "American Indian or Alaska Native" ~ "AIAN",
                     racerank1 == "Guamanian or Chamorro" | racerank1 == "Samoan" |
                       racerank1 == "Other Pacific Islander" ~ "NHPI",
                      racerank1 == "Some other race" ~ "Other",
                     .default = "missing")) %>% 
  filter(racecombo != "missing") %>% 
  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")

@@

>>>>

library(tidyverse)
library(dplyr)
library(tidyr)
library(here)
library(rio)
library(janitor)
library(knitr)
library(kableExtra)
library(mice)
```

@@

gss <- import(here("data", "GSS.xlsx")) %>%
  as_tibble() %>% 
  characterize() %>% # might not need characterize, TBD
  clean_names()

@@

```{r}
#| label: select-variables
data <- gss2022 %>% 
    select(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)

@@

```{r}
#| label: coding all missing data as NA 
# NOT WORKING - Need to troubleshoot
missing_vec <- c(".i:  Inapplicable" , ".n: No answer", ".d: Do not Know/Cannot Choose", ".s: Skipped on Web")

no_na_data <- data %>% 
  mutate(across(c(id, age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), ~ifelse(. %in% missing_vec, NA_character_, .)))

#data_complete <- mice(data, m=5)

#summary(complete(data,1))

@@

```{r}
#| label: mice imputations to deal with missing data

#data_complete <- mice(no_na_data, m=5)

#summary(complete(data_complete,1))
```

```{r}
#| label: recode-variables
tidydata <- data %>% 
  mutate(fear = 
           case_when( fear == ".i:  Inapplicable" | fear == ".n: No answer" | 
                        fear == ".d: Do not Know/Cannot Choose" |fear == ".s: Skipped on Web" ~ NA_character_, 
                      fear == "YES" ~ "yes", 
                      fear == "NO" ~ "no")) %>% 
  filter(fear == "yes" | fear == "no") %>% 
  mutate(racecombo = 
           case_when(racerank1 == "White" ~ "White",
                     racerank1 == "Black or African American" ~ "Black",
                     racerank1 == "Hispanic" ~ "Hispanic",
                     racerank1 == "Asian Indian" | racerank1 ==  "Chinese" |
                       racerank1 ==  "Filipino" | racerank1 ==  "Japanese" | 
                       racerank1 ==  "Korean" |  racerank1 ==  "Vietnamese" |
                       racerank1 ==  "Other Asian" ~ "Asian",
                     racerank1 == "American Indian or Alaska Native" ~ "AIAN",
                     racerank1 == "Guamanian or Chamorro" | racerank1 == "Samoan" |
                       racerank1 == "Other Pacific Islander" ~ "NHPI",
                      racerank1 == "Some other race" ~ "Other",
                     .default = NA_character_)) %>% 
  filter(!is.na(racecombo)) %>% 
  select("id", "age", "sex", "sexnow1", "sexornt", "degree", "attend", "raclive", "happy", "health", "helpful", "fair", "trust", "fear",  "partnrs5", "racecombo", "uscitzn", "dwelown")
